Global phase diagram and possible quantum spin liquid in the triangular $J-K-\Gamma$ model

### 2019-12-31 paper/paper.tex
20191231 18:23:40

\subsection{Moment direction of the classcial FM order}
For the classical FM order, the energy of the $J-K-\Gamma$ model per lattice site is given by
\begin{equation}
    E_{FM}^{c} = (3J + K) + 2\Gamma(S^xS^y + S^yS^z + S^zS^x)
    \label{eq:EcFM}
\end{equation}
On this level, the moment direction of the classical FM state is determined solely by $\Gamma$ and the problem becomes finding the global minimum and maximum of the multi-variable function $f(S^x, S^y, S^z) = S^xS^y + S^yS^z + S^zS^x$ with the constraint $|\mathbf{S}| = 1$. $f(S^x, S^y, S^z)$ takes maximum value $f_{max}=1$ at $S^x=S^y=S^z=\pm 1/\sqrt{3}$ and minimum value $f_{min}=-0.5$ when the conditions $S^x + S^y + S^z = 0$ and $|\mathbf{S}| = 1$ are fulfilled. The condition $S^x + S^y + S^z = 0$ specify a plane that is perpendicular to the $[111]$ direction, and considering the reference frame shown in Fig. \ref{fig:ModelDefinition}(c) this is actually the plane of the triangular lattice. That is to say, for $\Gamma > 0$ the moment of the classcial FM state prefers to lie in the lattice plane and when $\Gamma$ is negative, the moment would perpendicular to the lattice plane.

Our classical Monte Carlo simulation found FM ordered state with the right moment direction for majority of the area marked as FM phase (see Fig. 2 in Ref \onlinecite{PhysRevB.92.165108}) except for the pure $\Gamma$ term. For pure positive $\Gamma$ interaction, apart from the FM order lie in the lattice plane, we also found disordered states which have the same energy as the FM state but no states with lower energy were found. When $\Gamma=-1$, in addition to the FM order perpendicular to the lattice plane, other long-range ordered states which are energetically degenerate with the FM state appear (see Appendix A for details). Based on these observation, it is reasonable to say that the ground state of the classical $\Gamma$ model is highly degenerate though the degeneracy may be lifted by introducing other interactions. For example, a ferromagnetic ($J<0$) Heisenberg interaction would break the degeneracy and select the FM order as the ground state which was verified by our Monte Carlo simulation. On the other hand, when both positive Heisenberg and Kitaev interactions are included, which introduces further frustation to the system, the FM order is unlikely to be selected. For $\Gamma < 0$, the spins form stripe order and for positive $\Gamma$ the system remains disordered.

\subsection{Moment direction of the classcial stripe order}

In the case of stripe order, there are three degenerate spin configurations as illustrated in Fig. \ref{fig:PhaseDiagram}(b)-(d). For the spin configuration shown in Fig. \ref{fig:PhaseDiagram}(b), the energy of the model Hamiltonian \eqref{eq:Hamiltonian} is
\begin{IEEEeqnarray}{rCl}
    E_{StripeX}^{c} & = & -(J + K) + 2 K S^x S^x \nonumber \\
        && \negmedspace {} + 2 \Gamma (S^y S^z - S^z S^x - S^x S^y)
    \label{eq:EcStripeX}
\end{IEEEeqnarray}
and the energy for spin configuration in Fig. \ref{fig:PhaseDiagram}(c) and \ref{fig:PhaseDiagram}(d) can be obtained by a cyclic permutation $x \rightarrow y \rightarrow z \rightarrow x$. The moment direction of the classical stripe order is determined by the anisotropy parameters $K$ and $\Gamma$. In general, $E_{StripeX}^{c}$ has three extreme points where the first derivatives equal zero
\begin{align}
    &\mathbf{S}_0: \quad S_{0}^{y}=-S_{0}^{z} = \pm1/\sqrt{2}, \quad S_{0}^{x} = 0\\
    &\mathbf{S}_1: \quad S_{1}^{y}=S_{1}^{z} = f_{1}(K, \Gamma), \quad S_{1}^{x} = g_{1}(K, \Gamma) S_{1}^{y}\\
    &\mathbf{S}_2: \quad S_{2}^{y}=S_{2}^{z} = f_{2}(K, \Gamma), \quad S_{2}^{x} = g_{2}(K, \Gamma) S_{2}^{y}
\end{align}
see Appendix A for their explicit expressions.

For $\Gamma=0, K>0$, it can be clearly seen from Eq\eqref{eq:EcStripeX} that the stripe order shown in Fig. \ref{fig:PhaseDiagram}(b) prefers to lie in the $yz$ plane. A definite positive $\Gamma$ would fix the ordered moment to the $\mathbf{S}_0$ direction whereas negative $\Gamma$ drives it to the $\mathbf{S}_1$ direction. When $\Gamma=0, K<0$, the moment would along the $x$ axis to have lowest energy, both positive and negative $\Gamma$ make it deviate from the $x$ axis and point the $\mathbf{S}_1$ direction.

@@

To identify the remaining phases and providing more information of these phase, we calculate the SMSF for some representative model paramters, as shown in Fig. \ref{fig:StructureFactors}. For pure $\Gamma$ model, see Fig. \ref{fig:StructureFactors}(a) and (b), the peak of the SMSF locates at the center of first Brillouin Zone (BZ) which is a typical characteristic of the FM order.
Take into account of the classical results, we propose the three blue areas with different shadings to be FM ordered phases and they different from each other in the direction of the ordered moment which will be discussed in the next section. In Fig. \ref{fig:StructureFactors}(c)-(e), the SMSF peaked at the M-points indicating that the GS of the $J-K-\Gamma$ model around these model parameters could be stripe ordered.

>>>>

\subsection{Moment direction of the classcial FM order}
For the classical FM order, the energy of the $J-K-\Gamma$ model per lattice site is given by
\begin{equation}
    E_{FM}^{c} = (3J + K) + 2\Gamma(S^xS^y + S^yS^z + S^zS^x)
    \label{eq:EcFM}
\end{equation}
where $S^x$, $S^y$, $S^z$ are the corresponding components of the classical moment vector. On this level, the moment direction of the classical FM state is determined solely by $\Gamma$ and the problem becomes finding the global minimum and maximum of the multi-variable function $f(S^x, S^y, S^z) = S^xS^y + S^yS^z + S^zS^x$ with the constraint $|\mathbf{S}| = 1$. $f(S^x, S^y, S^z)$ takes maximum value $f_{max}=1$ at $S^x=S^y=S^z=\pm 1/\sqrt{3}$ and minimum value $f_{min}=-0.5$ when the conditions $S^x + S^y + S^z = 0$ and $|\mathbf{S}| = 1$ are fulfilled. The condition $S^x + S^y + S^z = 0$ specify a plane that is perpendicular to the $[111]$ direction, and considering the reference frame shown in Fig. \ref{fig:ModelDefinition}(c) this is actually the plane of the triangular lattice. That is to say, for $\Gamma > 0$ the moment of the classcial FM state prefers to lie in the lattice plane and when $\Gamma$ is negative, the moment would perpendicular to the lattice plane.

Our classical Monte Carlo simulation found FM ordered state with the right moment direction for majority of the area marked as FM phase (see Fig. 2 in Ref \onlinecite{PhysRevB.92.165108}) except for the pure $\Gamma$ term. For pure positive $\Gamma$ interaction, apart from the FM order lie in the lattice plane, we also found disordered states which have the same energy as the FM state but no states with lower energy were found. When $\Gamma=-1$, in addition to the FM order perpendicular to the lattice plane, other long-range ordered states which are energetically degenerate with the FM state appear (see Appendix A for details). Based on these observation, it is reasonable to say that the ground state of the classical $\Gamma$ model is highly degenerate though the degeneracy may be lifted by introducing other interactions. For example, a ferromagnetic ($J<0$) Heisenberg interaction would break the degeneracy and select the FM order as the ground state which was verified by our Monte Carlo simulation. On the other hand, when positive Heisenberg interaction is included, which introduces further frustation to the system, the FM order is unlikely to be the GS. In fact, the classical GS were found to be disordered for $0 < J \ll \Gamma$ and stripe ordered for $0 < J \ll -\Gamma$. As for the disordered phase, when quantum flucations are involved, a quantum spin liquid state may emerge.

\subsection{Moment direction of the classcial stripe order}
In the case of stripe order, there are three degenerate spin configurations as illustrated in Fig. \ref{fig:PhaseDiagram}(b)-(d). For the spin configuration shown in Fig. \ref{fig:PhaseDiagram}(b), the energy of the model Hamiltonian \eqref{eq:Hamiltonian} is
\begin{IEEEeqnarray}{rCl}
    E_{StripeX}^{c} & = & -(J + K) + 2 K S^x S^x \nonumber \\
        && \negmedspace {} + 2 \Gamma (S^y S^z - S^z S^x - S^x S^y)
    \label{eq:EcStripeX}
\end{IEEEeqnarray}
and the energy for spin configuration in Fig. \ref{fig:PhaseDiagram}(c) and \ref{fig:PhaseDiagram}(d) can be obtained by a cyclic permutation $x \rightarrow y \rightarrow z \rightarrow x$. The moment direction of the classical stripe order is determined by the anisotropy parameters $K$ and $\Gamma$. In general, $E_{StripeX}^{c}$ has three extreme points where the first derivatives equal zero
\begin{align}
    &\mathbf{S}_0: \quad S_{0}^{y}=-S_{0}^{z} = \pm1/\sqrt{2}, \quad S_{0}^{x} = 0 \label{eq:S0} \\
    &\mathbf{S}_1: \quad S_{1}^{y}=S_{1}^{z} = f_{1}(K, \Gamma), \quad S_{1}^{x} = g_{1}(K, \Gamma) S_{1}^{y} \label{eq:S1} \\
    &\mathbf{S}_2: \quad S_{2}^{y}=S_{2}^{z} = f_{2}(K, \Gamma), \quad S_{2}^{x} = g_{2}(K, \Gamma) S_{2}^{y} \label{eq:S2}
\end{align}
see Appendix B for their explicit expressions.

For $\Gamma=0, K>0$, it can be clearly seen from Eq\eqref{eq:EcStripeX} that the stripe order shown in Fig. \ref{fig:PhaseDiagram}(b) prefers to lie in the $yz$ plane. An infinitesimal positive $\Gamma$ would fix the ordered moment to the $\mathbf{S}_0$ direction whereas negative $\Gamma$ drives it to the $\mathbf{S}_1$ direction. When $\Gamma=0, K<0$, the moment would along the $x$ axis to have lowest energy, both positive and negative $\Gamma$ make it deviate from the $x$ axis to point the $\mathbf{S}_1$ direction. All these stripe orders with the specific moment direction were found in our Monte Carlo simulation.
However, in the HK limit ($\Gamma=0$), apart from the stripe orders lie in the axis plane, nematic orders which have the same energy were also found for a range of $J$ and $K$. Both positive and negative $\Gamma$ lift the degeneracy and drive the moment direction of stripe order to specific direction.
\textcolor{red}{$\cdots \cdots$}

@@

To identify the remaining phases and providing more information of these phases, we calculate the SMSF for some representative model paramters, as shown in Fig. \ref{fig:StructureFactors}. For pure $\Gamma$ model, see Fig. \ref{fig:StructureFactors}(a) and (b), the peak of the SMSF locates at the center of first Brillouin Zone (BZ) which is a typical characteristic of the FM order.
Take into account of the classical results, we propose the three blue areas with different shadings to be FM ordered phases and they different from each other in the direction of the ordered moment which will be discussed in the next section. In Fig. \ref{fig:StructureFactors}(c)-(e), the SMSF peaked at the M-points indicating that the GS of the $J-K-\Gamma$ model around these model parameters could be stripe ordered.
